---
title: "LINX"
author: "Peter Diakumis"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    css: style.css
---

```{r setup, message=FALSE}
require(gpgr)
require(devtools)
require(dplyr)
require(knitr)
require(DT)
```

```{r funcs}
dt_table <- function(x) {
  DT::datatable(
    x,
    filter = list(position = "top", clear = FALSE, plain = TRUE),
    class = "cell-border display compact",
    rownames = FALSE,
    extensions = c("Scroller", "Buttons", "KeyTable"),
    options = list(
      scroller = TRUE, scrollY = 400, scrollX = TRUE,
      autoWidth = FALSE, keys = TRUE,
      buttons = c("csv", "excel"), dom = "Blfrtip"
    )
  )
}
```


LINX
====
- <https://github.com/hartwigmedical/hmftools/tree/master/linx>.
- Annotates, interprets and visualises SVs.
  - groups SV calls together into events
- Inputs: GRIDSS SVs, PURPLE CNVs, gene panel, known fusions/fragile sites/LINEs

Table Outputs
-------------

### Annotations

- additional annotations of each non PON filtered break junction
- `svs.tsv`

```{r}
system.file("extdata/linx/tables/linx.svs.tsv", package = "gpgr") |>
  gpgr::linx_svs_read() |>
  dt_table()
```


### Breakends

- impact of each non PON filtered break junction on each overlapping gene
- `breakend.tsv`

```{r}
system.file("extdata/linx/tables/linx.breakend.tsv", package = "gpgr") |>
  gpgr::linx_breakend_read() |>
  dt_table()
```

### Clusters

- clustering of all non PON filtered SV events and their resolved
  classification.
- `clusters.tsv`

```{r}
system.file("extdata/linx/tables/linx.clusters.tsv", package = "gpgr") |>
  gpgr::linx_clusters_read() |>
  dt_table()
```

### Links

- segments joining break junction pairs predicted to be linked and phased in
  _cis_ on the derivative chromosome.
- `links.tsv`

```{r}
system.file("extdata/linx/tables/linx.links.tsv", package = "gpgr") |>
  gpgr::linx_links_read() |>
  dt_table()
```

### Fusions

- All inframe and outframe fusions predicted in the sample including HMF
  fusion knowledgebase annotations.
- `fusion.tsv`

```{r}
system.file("extdata/linx/tables/linx.fusion.tsv", package = "gpgr") |>
  gpgr::linx_fusion_read() |>
  dt_table()
```

### Driver Catalog

- Reproduction of the driver catalog produced by PURPLE with homozygous
  disruption events appended.
- `driver.catalog.tsv`

```{r}
system.file("extdata/linx/tables/linx.driver.catalog.tsv", package = "gpgr") |>
  gpgr::linx_drivercatalog_read() |>
  dt_table()
```

### Drivers

- Linkage of drivers from driver catalog to SV cluster which contributed to
  those drivers including LOH, deletion, disruption and amplification events.
- `drivers.tsv`

```{r}
system.file("extdata/linx/tables/linx.drivers.tsv", package = "gpgr") |>
  gpgr::linx_drivers_read() |>
  dt_table()
```

### Viral Inserts (Not Documented)

## Table 'vis' outputs

### Copy Number

- `linx.vis_copy_number.tsv`

```{r}
system.file("extdata/linx/tables/linx.vis_copy_number.tsv", package = "gpgr") |>
  gpgr::linx_viscopynumber_read() |>
  dt_table()
```

### Fusion

- `linx.vis_fusion.tsv`

```{r}
system.file("extdata/linx/tables/linx.vis_fusion.tsv", package = "gpgr") |>
  gpgr::linx_visfusion_read() |>
  dt_table()
```

### Gene Exon

- `linx.vis_gene_exon.tsv`

```{r}
system.file("extdata/linx/tables/linx.vis_gene_exon.tsv", package = "gpgr") |>
  gpgr::linx_visgeneexon_read() |>
  dt_table()
```

### Protein Domain

- `linx.vis_protein_domain.tsv`

```{r}
system.file("extdata/linx/tables/linx.vis_protein_domain.tsv", package = "gpgr") |>
  gpgr::linx_visproteindomain_read() |>
  dt_table()
```

### Segments

- `linx.vis_segments.tsv`

```{r}
system.file("extdata/linx/tables/linx.vis_segments.tsv", package = "gpgr") |>
  gpgr::linx_vissegments_read() |>
  dt_table()
```

### SV Data

- `linx.vis_sv_data.tsv`

```{r}
system.file("extdata/linx/tables/linx.vis_sv_data.tsv", package = "gpgr") |>
  gpgr::linx_vissvdata_read() |>
  dt_table()
```

## Session Info

```{r session_info, echo=FALSE}
pkgs_of_interest <- c("base", "gpgr")
si <- gpgr::session_info_kable(pkgs_of_interest)
si$si_pkg
si$si_pl
```
