---
title: "PURPLE"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
---

```{r knitr_opts, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
require(gpgr)
require(devtools)
require(dplyr)
require(knitr)
```

## Introduction

PURPLE: Purity Ploidy Estimator
(https://github.com/hartwigmedical/hmftools/tree/master/purple).

PURPLE combines B-allele frequency, read depth ratios, small variants and structural variants to
estimate the purity and copy number profile of a tumor sample.

It outputs several files, some of which are displayed below.

## Data Munging

### Somatic CNVs (per chromosome)

```{r cnv_som}
cnv_som <- system.file("extdata/purple/purple.cnv.somatic.tsv", package = "gpgr") |>
  gpgr::purple_cnv_som_process()
```

<details>
<summary>Description</summary>

```{r cnv_som_descr}
cnv_som$descr |>
  knitr::kable(caption = "PURPLE Somatic CNVs (per chromosome) Columns.")
```

</details>

```{r cnv_som_tab}
cnv_som$tab |>
  dplyr::slice(1:10) |>
  knitr::kable(caption = "PURPLE Somatic CNVs (per chromosome) Summary Table.")
```

### Somatic CNVs (per gene)

```{r cnv_som_gene}
umccr_key_genes <- system.file("extdata/ref/somatic_panel-v24.03.0.tsv", package = "gpgr")
cnv_som_gene <- system.file("extdata/purple/purple.cnv.gene.tsv", package = "gpgr") |>
  gpgr::purple_cnv_som_gene_process(g = umccr_key_genes)
```

<details>
<summary>Description</summary>

```{r cnv_som_gene_descr}
cnv_som_gene$descr |>
  knitr::kable(caption = "PURPLE Somatic CNVs (per gene) Columns.")
```

</details>

```{r cnv_som_gene_tab}
cnv_som_gene$tab |>
  dplyr::slice(1:10) |>
  knitr::kable(caption = "PURPLE Somatic CNVs (per gene) Summary Table.")
```

### Purity

```{r purity}
purity <- system.file("extdata/purple/purple.purity.tsv", package = "gpgr") |>
  gpgr::purple_purity_read()

purity$summary |>
  knitr::kable(caption = "PURPLE Purity Summary Table.")
```

### Kataegis

```{r kataegis}
kat <- system.file("extdata/purple/purple.somatic.vcf.gz", package = "gpgr") |>
  purple_kataegis()
kat$data |>
  dplyr::slice(1:10) |>
  knitr::kable(caption = "PURPLE Kataegis Table.")
```

<details>
<summary>Description</summary>

```{r kataegis_descr}
knitr::kable(kat$description, caption = "Kataegis column description.")
```

</details>

### QC

```{r qc}
qc <- system.file("extdata/purple/purple.qc", package = "gpgr") |>
  gpgr::purple_qc_read()

qc$summary |>
  knitr::kable(caption = "PURPLE QC Summary Table.")
```

## Session Info

```{r session_info, echo=FALSE}
pkgs_of_interest <- c("base", "gpgr")
si <- gpgr::session_info_kable(pkgs_of_interest)
si$si_pkg
si$si_pl
```
